---
title: "dplyr in ten minutes!"
author: "Corinne Riddell (post-doc @McGill)"
date: "3/15/2018"
output: md_document
---

##What is dplyr?

- wrangling/cleaning/manipulation
- human readable
- best when paired with the piping operator %>%

##Install and load dplyr

```{r}
#install.packages("dplyr")
library(dplyr)
```

##Example data

- We'll use a dataset on life expectancy for black and white Americans to practice using `dplyr`
- Let's start by importing these data into R

```{r import-data}
# Read in the dataset
BlackWhite_results <- read.csv("./BlackWhite_results.csv")
```

##Take a peak

- Let's use RStudio's viewer to look at these data.
- Click the table icon next to BlackWhite_results in the environment tab or run `View(BlackWhite_results)`
- If you aren't using RStudio, you can use `head(BlackWhite_results)` to see the first six lines of data

```{r dimensions-and-names}
dim(BlackWhite_results)
names(BlackWhite_results)
```

## So. many. variables.

Let's only keep a subset for further use.

##dplyr select() to keep or discard variables {.build}

- `select()` is used to select a subset of variables from an existing data frame.
- You can explicitly select the variables you'd like to keep, or "negatively" 
`select()` (i.e, remove) variables you'd like to discard.

```{r select-example}
subset <- select(BlackWhite_results, 
                 LE_white_mean, LE_black_mean,
                 state, year, sex, stabbrs)

names(subset)
dim(subset)
```

##dplyr select() to keep or discard variables (columns) {.build}

Write this another way using the **piping** operator (%>%)

```{r select-pipe-example}
subset2 <- BlackWhite_results %>% 
  select(LE_white_mean, LE_black_mean,
         state, year, sex, stabbrs)
```

Confirm that `subset` and `subset2` are identical:
```{r check-identical}
identical(subset, subset2)

BlackWhite_results <- subset2
```

## dplyr mutate() to add a new variable {.build}

- What if you wanted to calculate the difference (in years) between the life expectancy
for blacks and whites?

```{r mutate-check}
BlackWhite_results <- BlackWhite_results %>% 
  mutate(diff = LE_white_mean - LE_black_mean)
```

## dplyr mutate() to add a new variable {.build}

- Take a peak to see the calculated difference

```{r view-results}
BlackWhite_results %>% select(LE_white_mean, LE_black_mean, 
                              diff, state, year, sex) 
```
                              
## dplyr rename() to rename a variable {.build}
- Let's rename the `sex` variable as `gender`

```{r solution-ex4}
BlackWhite_results <- BlackWhite_results %>% 
  rename(gender = sex)

names(BlackWhite_results)
```

## dplyr arrange() to sort the dataset based on one or several variables {.build}
- Try `arrange()`ing the dataset by the `diff` variable, and by `-diff`
- Have a look at the results to compare

```{r solution-ex5}
Results_sort1 <- BlackWhite_results %>% 
  arrange(diff)

Results_sort2 <- BlackWhite_results %>% 
  arrange(-diff)
```

## dplyr filter() to subset the data's rows based on a condition {.build}
- Let's use `filter()` to create a new dataset `ny.women` that contains only women from New York.

```{r filter}
ny.women <- BlackWhite_results %>% 
  filter(state == "New York", gender == "Female")

dim(ny.women)
```

## dplyr summarise() to perform an aggregated calculation
- Summarise can be used to calculate summary statistics
- What variable does this code summarize over?

```{r summarise}
ny.women %>% 
  summarise(average.le.white = round(mean(LE_white_mean), 1),
            average.le.black = round(mean(LE_black_mean), 1))
```

## Stacking dplyr() functions using %>%
- What does this code do?
```{r filter-and-summarise-stacked}
 BlackWhite_results %>% 
  filter(state == "New York", gender == "Female", year > 2010) %>%
  summarise(average.le.white = round(mean(LE_white_mean), 1),
            average.le.black = round(mean(LE_black_mean), 1))
```

## dplyr group_by() %>% summarise() 
- Pair together `group_by()` and `summarise()` to calculate summary statistics within groups

```{r}
yearly.means <- BlackWhite_results %>% 
  group_by(year, gender) %>%
  summarise(ave.yr.gender = mean(diff))
```

## dplyr cheat sheet

- https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf

## Connect
  
- Mail: corinne.riddell@mail.mcgill.ca
- Web: corinneriddell.com
- GitHub: corinne-riddell
- Twitter: @datavisitor
