---
title: "Mini Introductory Workshop to SNA"
author: "Jennifer Yu"
date: "November 3, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = TRUE,
	warning = TRUE
)
```

# Description 

In this mini workshop to social network analysis, we will learn on how to get data into R, to illustrate a network graph and to calculate some network measures.

# Installing packages 
```{r}
install.packages("igraph") # install the igraph package

library(igraph) # load the igraph package

```

# Setting up the working directory 
```{r}
setwd("C:/Users/Jennifer Yu/RLadies/SNA") # Change for the directory where you saved the files

```

# Importing databases 
```{r}
nodes <- read.csv("Rladies_SNA_nodes.csv", stringsAsFactors = F) # database containing the nodes' attributes 
adj.matrix <- read.csv("Rladies_SNA_edges.csv", stringsAsFactors = F) # database containing the adjacency matrix, i.e. the connections between the nodes 

```

# Explore databases 
```{r}
nodes # database with the characteristics of the nodes, the attendees of last meetup of R-ladies

adj.matrix # database on the edges (ties) of the network 

```

# Creating the network graph 

## Create the backbone of the network graph with the data on edges (ties)
```{r}
class(adj.matrix) # Check whether our adjacency matrix is a matrix

adj.mat  <- as.matrix(adj.matrix[,-1]) # coerces the data set as a matrix if not already a matrix
rownames(adj.mat) <- colnames(adj.mat)

rladies.graph <- graph.adjacency(adj.mat, mode="directed", weighted = TRUE, diag = FALSE) # this will create an 'igraph object'

rladies.graph # Explore the network graph

summary(rladies.graph) # Let's say we just want to see a small summary of the network graph, and we don't want to see all the ties

V(rladies.graph) # Look at the vertices/nodes

E(rladies.graph) # Look at the edges/ties

```

## Add node attributes 
```{r}
vertex_attr(rladies.graph, "ID", index = V(rladies.graph)) <- nodes$ID # Add the ID label to all nodes 

vertex_attr(rladies.graph, "Season", index = V(rladies.graph)) <- nodes$season # Add the favorite season for each node

vertex_attr(rladies.graph, "Dining", index = V(rladies.graph)) <- nodes$restaurant # Add the number of times went out dining for each node

vertex_attr(rladies.graph, "Personality", index = V(rladies.graph)) <- nodes$extroversion # Add the personality type for each node

rladies.graph # Take a look at the network graph now

```

# Plotting the network graph 

## Add some aesthetic features to the network graph
```{r}
rladies.plot <- rladies.graph # Create a new graph just for the plotting, unnecessary step, just to see things after

V(rladies.plot)[V(rladies.plot)$Season == "spring"]$color <- "pink" # Add the colour for the season
V(rladies.plot)[V(rladies.plot)$Season == "summer"]$color <- "green3"
V(rladies.plot)[V(rladies.plot)$Season == "fall"]$color <- "orange"
V(rladies.plot)[V(rladies.plot)$Season == "winter"]$color <- "skyblue"
V(rladies.plot)[V(rladies.plot)$Season == "missing"]$color <- "grey50"

V(rladies.plot)$frame.color <- NA # remove frame color of nodes

V(rladies.plot)$size <- V(rladies.graph)$Dining # Add the size of the nodes for number of times dinning out


V(rladies.plot)[V(rladies.plot)$Personality == "extrovert"]$shape <- "star" # Add the node shape for personality type
V(rladies.plot)[V(rladies.plot)$Personality == "ambivert"]$shape <- "circle"
V(rladies.plot)[V(rladies.plot)$Personality == "introvert"]$shape <- "square"
V(rladies.plot)[V(rladies.plot)$Personality == "missing"]$shape <- "circle"

V(rladies.plot)$label <- NA #remove node labelling


E(rladies.plot)$width <- E(rladies.graph)$weight^2 # add edge width

E(rladies.plot)$arrow.size <- 0.5 # specify arrow size

```

## Plotting! 
```{r}
plot(rladies.graph) # plotting the graph without any aesthetics

plot(rladies.plot) # Simple plotting with aesthetics

plot(rladies.plot, layout = layout.circle) # circle layout

plot(rladies.plot, layout = layout.fruchterman.reingold) # another layout

```


# Calculating some measures  
From the summary of the network graph, we can see the network size (number of nodes) and the number of edges (ties)


## Network density
```{r}

edge_density(rladies.graph, loops = FALSE) # we don't have loops in the network, so FALSE for loops

```

## Degree centrality 
```{r}
centr_degree(rladies.graph, mode = c("all"), loops = FALSE, normalized = FALSE) # I want all the ties, I don't have loops and I don't want the proportion 

centr_degree(rladies.graph, mode = c("out"), loops = FALSE, normalized = FALSE) # Here the number of people that each person had nominated 

centr_degree(rladies.graph, mode = c("in"), loops = FALSE, normalized = FALSE) # Here the number of times each person has been nominated

```

There are many things to play with, so feel free to explore!




